---
title: "STA 108 Project 1"
author: "Darren Lam and Andy Kim"
date: "2024-10-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align =
"center", fig.width = 10)
options(scipen = 999) #Remove the scientific notation

# Load libraries
library(ggplot2)

# Import the dataset
Senic <- read.csv("SENIC.csv")

```

```{r Part 2, echo=FALSE}
# Summary stats for each variable
summary(Senic$length)
summary(Senic$infection)
summary(Senic$facility)
summary(Senic$Xray)

```

```{r Part 3, echo=FALSE}
# Scatter Plot of Infection vs. Length
ggplot(data = Senic, mapping = aes(x = infection, y = length)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot of Infection Probability vs. Length of Hospital Stay",
       x = "Probability of Acquiring Infection (%)",
       y = "Length of Hospital Stay (Days)")

# Summary Table and Model of Infection vs. Length
infection_model <- lm(length ~ infection, data = Senic)
summary(infection_model)

# Scatter Plot of Facility vs. Length
ggplot(data = Senic, mapping = aes(x = facility, y = length)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot of Facilities and Services Provided vs. Length of Hospital Stay",
       x = "% of 35 Potential Facilities and Services Provided",
       y = "Length of Hospital Stay (Days)")

# Summary Table and Model of Facility vs. Length
facility_model<- lm(length ~ facility, data = Senic)
summary(facility_model)

# Scatter Plot of Xray vs. Length
ggplot(data = Senic, mapping = aes(x = Xray, y = length)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot of X-ray Ratio vs. Length of Hospital Stay",
       x = "Ratio of X-rays Performed to # of Patients Without Signs of Pneumonia",
       y = "Length of Hospital Stay (Days)")

# Summary Table and Model of Xray vs. Length
Xray_model <- lm(length ~ Xray, data = Senic)
summary(Xray_model)

```

```{r Part 4, echo=FALSE}
# Residual plot
ggplot(data = infection_model, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  labs(title = "Plot of Residuals for Infection Probability vs. Length of Hospital Stay",
       x = "Probability of Acquiring Infection (%)",
       y = "Residuals")

# Histogram of residuals
ggplot(data = infection_model, mapping = aes(x = .resid)) +
  geom_histogram() +
  labs(title = "Histogram of Residuals for Infection Probability vs. Length of Hospital Stay",
       x = "Residuals",
       y = "Frequency")

# QQ plot
ggplot(data = infection_model, mapping = aes(sample = .resid)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Normal QQ plot for Infection Probability vs. Length of Hospital Stay",
       x = "Theoretical Quantities",
       y = "Sample Quanitities")

```

Our final simple linear regression model used the infection probability as a predictor to explain the length of hospital stays with the highest degree of fit. The Shapiro-Wilks test (for normality) and the Fligner-Killeen test (for constant variance) were used for the analysis portion of the report. For the Shapiro-wilks test the null hypothesis was that the population is normally distributed. The alternative hypothesis was that the population was not normally distributed. The p-value was 0.00000001699 which is less than the alpha of 0.05 so we rejected the null hypothesis and concluded that the population is not normally distributed. For the Fligner-Killeen test, the null hypothesis was that the population has a constant variance. The alternative hypothesis was that the population did not have a constant variance. The p-value was 0.2543 which is greater than the alpha of 0.05 so we had no evidence to reject the null hypothesis and concluded that the model suggests that the population has a constant variance. These contradictions could be a result of the outliers in the data. R squared is rounded to 0.285 so the model explains 28.5% of the variation in the length of stay. 


```{r Part 5, echo=FALSE}
# Confidence Interval for slope
confint(infection_model)[2, ]

# Shapiro-wilks test (for normality)
shapiro.test(infection_model$residuals)

# Fligner-killeen test (for constant variance)
groups <- rep("Lower", nrow(Senic))
groups[Senic$infection > median(Senic$infection)] = "Upper"
groups <- as.factor(groups)
Senic$group <- groups
fligner.test(infection_model$residuals, Senic$group)

```

```{r Part 7, echo=FALSE}
# Predictions
predict(infection_model, newdata = data.frame(infection = 5.2), interval = "confidence")
predict(infection_model, newdata = data.frame(infection = 4.4), interval = "prediction")
predict(infection_model, newdata = data.frame(infection = 3.7), interval = "confidence")

```

# R Appendix
```{r, ref.label=knitr::all_labels(), eval = F, echo = T}
tinytex::install_tinytex()
```